Theory about Normal Basis on $F(2^k)$

### 2016-12-11 dissertation/prelim_basis.tex
dissertation v3:milestone

\chapter{Gr\"ober Basis and Algebraic Geometry} \label{ch:ideals}
This chapter reviews fundamental concepts of commutative and 
computer algebra which are used in this work. 
Specifically, this chapter covers monomial ordering, polynomial ideals and 
varieties, and the computation of \Grobner bases.
It also overviews elimination theory as well as Hilbert's Nullstellensatz 
theorems and how
they apply to Galois fields. The results of these theorems are used in polynomial
abstraction and formal verification of Galois field circuits and 
are discussed 
in subsequent chapters. The material of this chapter is
mostly referred from the textbooks \cite{ideals:book,gb_book} and 
previous work by {\it Lv} \cite{lv:phd} as well as {\it Pruss} \cite{pruss:tcad15}.

@@

Thus, $x^2\cdot y$ is a monomial in variables $x,y$ with total degree $3$.
For simplicity, we will henceforth denote a 
monomial $x_1^{{\alpha}_1} \cdot x_2^
{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d}$ as $x^{\alpha}$, 
where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.

@@

A {\bf lexicographic ordering} (lex) is a total-ordering $>$ such that 
variables in the terms are lexicographically ordered, i.e. simply based on 
when the variables appear in the ordering.
Higher variable-degrees take 
precedence over lower degrees for equivalent variables (e.g. $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
\begin{Definition}
{\bf Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
& \text{Starting  from the  left, the first co-ordinates of $\alpha_i, \beta_i$} \\
& \text{that are different satisfy $\alpha_i > \beta_i$}

@@

A {\bf degree-lexicographic ordering} (deglex) is a total-ordering $>$ such 
that the total degree of a term takes precedence over the lexicographic 
ordering.  
A {\bf degree-reverse-lexicographic ordering} (degrevlex) is the same as a
deglex ordering, however terms are lexed in reverse.

\begin{Definition}
{\bf Degree Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
\sum_{i=1}^{d}\alpha_i > \sum_{i=1}^{d} \beta_i & \text{ or }\\
\sum_{i=1}^{d}\alpha_i = \sum_{i=1}^{d} \beta_i  \text{ and }
x^{\alpha} > x^{\beta} & \text{w.r.t. lex order}
\end{cases}
\end{equation}
\end{Definition}

@@

\begin{Definition}
The {\bf leading term} is the first term in a term-ordered polynomial.
Likewise, the {\bf leading coefficient} is the coefficient of
the leading term. 
Finally, a {\bf leading monomial} is the leading term 
lacking the coefficient.  We use the following notation:
\begin{eqnarray}
     lt(f)&& \text{--- Leading Term} \\
     lc(f)&& \text{--- Leading Coefficient} \\
     lm(f)&& \text{--- Leading Monomial} \\
     tail(f)&& f - lt(f)
\end{eqnarray}
\end{Definition}

@@

\begin{Definition}
Let $\F$ be a field and let $f, g \in \F[x_1, x_2, \ldots, x_d]$ be polynomials
over the field. {\bf Polynomial division} of $f$ by $g$ is computes following:
\begin{equation}
f-\frac{lt(f)}{lt(g)}\cdot g
\end{equation}
This polynomial division is denoted
\begin{equation}
f\xrightarrow{g} r
\end{equation}
where $r$ is the resulting polynomial of the division.
If $\frac{lt(f)}{lt(g)}$ is non-zero, then $f$ is considered divisible by $g$, 
i.e. $g \mid f$.
\end{Definition}
Notice that if $g \nmid f$, that is if $f$ is not divisible by $g$,
then the division operation gives $r = f$.
\begin{Example}
Over $\R[x,y,z]$, set the lex term order $x > y > z$.
Let $f = -2x^3 + 2x^2yz + 3xy^3$ and $g = x^2+yz$.
\begin{equation}
\frac{lt(f)}{lt(g)} = \frac{-2x^3}{x^2} = -2x
\end{equation}
Since $\frac{lt(f)}{lt(g)}$ is non-zero $g|f$. The division, $f\xrightarrow{g} r$, 
is computed as:
\begin{eqnarray}
r &=& f-\frac{lt(f)}{lt(g)}\cdot g = -2x^3 + 2x^2yz + 3xy^3 - (-2x \cdot (x^2+yz)) \nonumber \\
&=& -2x^3 + 2x^2yz + 3xy^3 - (-2x^3-2xyz) = 2x^2yz + 3xy^3 + 2xyz
\end{eqnarray}
Notice that the division $f\stackrel{g}{\textstyle\longrightarrow}r$ cancels the leading term of $f$.
\end{Example}

Similarly, we can also define when a polynomial is divided (reduced) by a set of polynomials.
\begin{Definition}
    The {\bf reduction} of a polynomial $f$, by another polynomial $g$, to
    a reduced polynomial $r$ is denoted:
    \begin{equation*}
        f\stackrel{g}{\textstyle\longrightarrow}_{+}r
    \end{equation*}
    which is transitive and reflective closure of relation $f\stackrel{g}{\textstyle\longrightarrow}r$. 
    Reduction is carried out using multivariate, polynomial long division. 
    % The long division is performed according to a term-ordering on polynomials, and the division algorithm 
    %terminates when the leading term of the divisor does not divide any 
    %other term in the dividend.

@@

The above example shows the variety can be infinite, finite (non-empty set) 
or empty. It is interesting to note that since we will be operating over
finite fields $\F_{q}$, and any finite set of points is a variety. Likewise,
any variety over $\F_{q}$ is finite (or empty).
Consider the points $\{(a_1,\dots, a_d): a_1, \dots, a_d \in \F_q\}$
in $\F_q^d$. Any single point is a variety of some polynomial system:
e.g. $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
varieties are also varieties.

@@

The same ideal may have many different bases.
For instance, it is possible to have different sets of polynomials
$\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
ideal, i.e., 
$\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
variety depends on the ideal, these sets of polynomials have the same 
solutions.

@@

Ideals and their varieties are a key part of computer-algebra based formal
verification. A given hardware design can be transformed into a set
of polynomials over a field, $f_1, \ldots, f_s \in F$.
This set of polynomials gives the system of equations:
\begin{eqnarray}
f_1 = 0 \nonumber \\
\vdots \nonumber \\
f_s = 0 \nonumber
\end{eqnarray}
Using algebra, it is possible to derive new equations from the original 
system.
The ideal $\langle f_1,\ldots, f_s \rangle$ provides a way of analyzing 
such {\it consequences} of a system of polynomials.

\begin{Example}
Given two equations in $\mathbb{R}[x,y,z]$:
\begin{eqnarray}
x=z+1 \nonumber \\
y=x^2+1 \nonumber 
\end{eqnarray}
we can eliminate $x$ to obtain a new equation:
\begin{equation}
y=(z+1)^2+1=z^2+2z+2 \nonumber 
\end{equation}
Let $f_1, f_2, h \in \mathbb{R}[x,y,z]$ be polynomials based on these 
equations:
\begin{eqnarray}
f_1 &= x-z-1 &= 0 \nonumber \\
f_2 &= y-x^2-1 &= 0 \nonumber \\
h   &= y-z^2-2z-2 &= 0 \nonumber
\end{eqnarray}
If $I$ is the ideal generated by $f_1$ and $f_2$, i.e. 
$I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
\begin{eqnarray}
g_1 &= x+z+1 \nonumber \\
g_2 &= 1     \nonumber \\
h &= g_1\cdot f_1+g_2\cdot f_2  = y-z^2-2z-2 \nonumber
\end{eqnarray}
where $g_1, g_2 \in \mathbb{R}[x,y,z]$.
Thus, we call $h$ a {\bf member of the ideal} $I$.
\end{Example}

@@

Let $\mathbb{F}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{F}^d$ be a point, and $f \in
\mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ if $f(\mathbf{a}) = 0$, i.e.,
$\mathbf{a}$ is in the variety of $f$.

@@

As mentioned earlier, different polynomial sets may generate the same 
ideal. Some of these generating sets may be a better representation of 
the ideal, and thus provide more information and insight into the properties 
of the ideal. One such ideal representation is a {\bf \Grobner basis}, which has
a number of important properties that can solve numerous polynomial 
decision questions:

\begin{itemize}
\item Presence or absence of solutions (varieties)
\item Dimension of the varieties
\item Ideal membership of a polynomial
\item $\dots \dots$
\end{itemize}

@@

\begin{Example}
Consider Example \ref{exp:gbsimple}. Let $f = y^2x - x$ be another
polynomial. Note that $f = yf_1 + f_2$, so $f \in I$. If we divide $f$
by $f_1$ first and then by $f_2$, we will obtain a zero
remainder. However, since the set $\{f_1, f_2\}$ is not a Gr\"{o}bner
basis, we find that the reduction $f
\stackrel{f_2}{\textstyle\longrightarrow} x^2 - x
\stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
i.e. dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
$I$, $G = \{x^2 - x, yx - y, y^2 - x\}$, dividing $f$ by polynomials
in $G$ in any order will always lead to the zero remainder. Therefore,
one can decide ideal membership unequivocally using the Gr\"{o}bner
basis. 
\end{Example}

@@

\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{$F = \{f_1, \dots, f_s\}$, such that $I=\langle f_1, \dots, f_s\rangle$}, and term order $>$
 \KwOut{$G = \{g_1,\dots ,g_t\}$, a Gr\"{o}bner basis of $I$ }
  $G:= F$\;
  \Repeat{$G = G'$}
  {
  	$G' := G$\;
  	\For{ each pair $\{f_{i}, f_{j}\}, i \neq j$ in $G'$} 
	{
		$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$ \;
		\If{$r \neq 0$}
		{
			$G:= G \cup \{r\}$ \;
		}
	}
   }
\caption {Buchberger's Algorithm}\label{alg:gb}
\end{algorithm}

@@

When computing a \Grobner basis, it's important to note that if $lt(f_i)$ 
and $lt(f_j)$ have no common variables, the S-poly reduction step in 
Buchberger's algorithm will reduce to 0.
\begin{Lemma}
In Buchberger's algorithm, when $gcd(lt(f_i),lt(f_j)) = 0$, the S-poly reduction
$$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$$
will produce $r=0$.
\end{Lemma}

@@

\begin{Definition}\label{def:minigb}
A {\bf minimal Gr\"obner basis} for a polynomial ideal $I$ is a \Grobner basis $G$ for $I$ such that
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$,  $lt(g_{i}) \notin \langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
A {\bf minimal} \Grobner basis is a \Grobner basis such that all polynomials
have a coefficient of $1$ and no leading term of any element in $G$ divides 
another in $G$.
Given a \Grobner basis $G$, a minimal \Grobner basis can be
computed as follows:
\begin{enumerate}
\item Minimize every $g_i \in G$, i.e $g_i=g_i/lc(g_i)$
\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, i.e. remove every polynomial in $G$ whose leading term is divisible by the leading term of some other polynomial in $G$.
\end{enumerate}

@@

\begin{Definition}
Let $H = \{h_1, \ldots, h_t\}$ be a minimal Gr\"obner basis.  Apply
the following reduction process: 
\begin{itemize}
\item $h_1 \stackrel{G_1}{\textstyle\longrightarrow}_+ g_1$, where
  $g_1$ is reduced w.r.t. $G_1 = \{h_2, \ldots, h_t\}$

\item $h_2 \stackrel{G_2}{\textstyle\longrightarrow}_+ g_2$, where
  $g_2$ is reduced w.r.t. $G_2 = \{g_1, h_3, \ldots, h_t\}$
\item $h_3 \stackrel{G_3}{\textstyle\longrightarrow}_+ g_3$, where
  $g_3$ is reduced w.r.t. $G_3 = \{g_1, g_2, h_4, \ldots, h_t\}$

\hspace{0.25in} $\vdots$
\vspace{0.1in}
\item $h_t \stackrel{G_t}{\textstyle\longrightarrow}_+ g_t$, where
  $g_t$ is reduced w.r.t. $G_t = \{g_1, g_2, g_3, \ldots, g_{t-1}\}$
\end{itemize}
Then $G = \{g_1, \ldots, g_t\}$ is a {\bf reduced Gr\"obner basis.}
\end{Definition}

@@

% TODO discuss complexity

The high computational complexity if Gr\"obner basis computation is treated as an issue
because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
the worst case computation time/space cost of its Gr\"obner basis is doubly exponential bounded -- $q^{q^{O(|\phi|)}}$
\cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
the polynomial is well restricted rather than arbitrary.
Gao et al. \cite{gao:qe-gf-gb} proves that when the number of variables and degree of polynomials 
are restricted, the complexity reduces to single exponential $q^{O(|\phi|)}$.
This provides the possibility to make use of Gr\"obner basis under restricted situations,
which is the theoretic basis of this dissertation.

\Grobner basis computation depends on the $Spoly$ computation, which in turn 
depends on the leading terms of polynomials. Thus, different monomial 
orderings can result in different \Grobner basis computations for the 
same ideal. Computation using a degrevlex ordering tends to be least 
difficult, while lex ordering tends to be computationally complex. However, 
lex ordering used in the computation of \Grobner basis is an {\bf elimination
ordering}; that is, the polynomials contained in the resulting \Grobner basis
have continuously eliminated variables in the ordering. This is the topic of 
elimination theory, which is described in the following sections as well as 
its theoretic basis -- the Nullstellensatz theory.

@@

%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
\begin{Theorem}{Weak Nullstellensatz}
%$\left[\bf{Weak\  Nullstellensatz}\right]$ 
Let $J \subset \overline {\F}[x_1, x_2, \cdots, x_d]$ 
be an ideal satisfying $V(J)=\emptyset$. 
Then $J=\overline {\F}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
\begin{equation}
V(J)=\emptyset\ \iff\ J=\overline {\F}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
\end{equation}
\end{Theorem}

@@

{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
system of multivariate polynomial equations (ideal $J$) has common solutions 
in ${\overline {\F}}^d$. For this purpose, we only need to check if 
the ideal is generated by the unit element, i.e., $1\in J$. 
This approach can be used to evaluate the feasibility of constraints in 
verification problems.
%%%%%%%%%%%%%%%%%%%%%%%%%strong Nullstellensatz%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
An interesting result is one of {\bf Strong 
Nullstellensatz}.
The strong
Nullstellensatz establishes the correspondence between radical ideals
and varieties.

\begin{Theorem}\label{thm:sns}
({\it The Strong Nullstellensatz} \cite{gb_book}) 
Let $\overline{\F}$ be an algebraically closed field, and let $J$
be an ideal in $\overline{\F}[x_1,\dots, x_d]$. 
Then we have $I(V_{\overline{\F}}(J)) =\sqrt{J}$. 
\end{Theorem}

@@

$g_8=Z+A\cdot B$ is the {\bf canonical, word-level polynomial } 
representing the function performed by the multiplier $Z=A\cdot B$.
\end{Example}

\section{Concluding Remarks}
Our approach to word-level abstraction of Galois field arithmetic 
circuits applies concepts of polynomial ideals, varieties, \Grobner basis, 
and abstraction theory to implement verifications on sequential circuits. 
These approaches are described in the following chapters.

>>>>

\chapter{Gr\"ober Bases and Algebraic Geometry} \label{ch:ideals}
This chapter reviews fundamental concepts of commutative and 
computer algebra which are used in this work. 
Specifically, this chapter covers monomial ordering, polynomial ideals and 
varieties, and the computation of \Grobner bases.
It also overviews elimination theory as well as Hilbert's Nullstellensatz 
theorems and how
they apply to Galois fields. The results of these theorems are used in polynomial
abstraction and formal verification of Galois field circuits and 
are discussed 
in subsequent chapters. The material of this chapter is
mostly referred from the textbooks \cite{ideals:book,gb_book} and 
previous work by Lv \cite{lv:phd} as well as Pruss \cite{pruss:tcad15}.

@@

Thus, $x^2\cdot y$ is a monomial in variables $x,y$ with total degree $3$.
For simplicity, we will henceforth denote a 
monomial $x_1^{{\alpha}_1} \cdot x_2^
{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d}$ as $x^{\alpha}$, 
where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
integers $\ge 0$, {\it i.e.} $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.

@@

A {\bf lexicographic ordering} (LEX) is a total-ordering $>$ such that 
variables in the terms are lexicographically ordered, {\it i.e.} simply based on 
when the variables appear in the ordering. It is also a well-order, where the least element exists.
Higher variable-degrees take 
precedence over lower degrees for equivalent variables ({\it e.g.} $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
\begin{Definition}
{\bf Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
& \text{Starting  from the  left, the first co-ordinates of $\alpha_i, \beta_i$} \\
& \text{that are different satisfy $\alpha_i > \beta_i$}

@@

A {\bf degree-lexicographic ordering} (DEGLEX) is a total-ordering $>$ such 
that the total degree of a term takes precedence over the lexicographic 
ordering.  
A {\bf degree-reverse-lexicographic ordering} (DEGREVLEX) is the same as a
DEGLEX ordering, however terms are LEXed in reverse.

\begin{Definition}
{\bf Degree Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
\sum_{i=1}^{d}\alpha_i > \sum_{i=1}^{d} \beta_i & \text{ or }\\
\sum_{i=1}^{d}\alpha_i = \sum_{i=1}^{d} \beta_i  \text{ and }
x^{\alpha} > x^{\beta} & \textit{w.r.t.}\text{ LEX order}
\end{cases}
\end{equation}
\end{Definition}

@@

\begin{Definition}
The {\bf leading term} is the first term in a term-ordered polynomial.
Likewise, the {\bf leading coefficient} is the coefficient of
the leading term. 
Finally, a {\bf leading monomial} is the leading term 
lacking the coefficient.  We use the following notation:
\begin{eqnarray}
     lt(f)&& \text{Leading Term} \\
     lc(f)&& \text{Leading Coefficient} \\
     lm(f)&& \text{Leading Monomial} \\
     tail(f)&& f - lt(f)
\end{eqnarray}
\end{Definition}

@@

\begin{Definition}
Let $\F$ be a field and let $f, g \in \F[x_1, x_2, \ldots, x_d]$ be polynomials
over the field. {\bf Polynomial division} of $f$ by $g$ is computes following:
\begin{equation}
f-\frac{lt(f)}{lt(g)}\cdot g
\end{equation}
This polynomial division is denoted
\begin{equation}
f\xrightarrow{g} r
\end{equation}
where $r$ is the resulting polynomial of the division.
If $\frac{lt(f)}{lt(g)}$ is non-zero, then $f$ is considered divisible by $g$, 
{\it i.e.} $g \mid f$.
\end{Definition}
Notice that if $g \nmid f$, that is if $f$ is not divisible by $g$,
then the division operation gives $r = f$.
\begin{Example}
Over $\R[x,y,z]$, set the lex term order $x > y > z$.
Let $f = -2x^3 + 2x^2yz + 3xy^3$ and $g = x^2+yz$.
\begin{equation}
\frac{lt(f)}{lt(g)} = \frac{-2x^3}{x^2} = -2x
\end{equation}
Since $\frac{lt(f)}{lt(g)}$ is non-zero $g|f$. The division, $f\xrightarrow{g} r$, 
is computed as:
\begin{eqnarray}
r &=& f-\frac{lt(f)}{lt(g)}\cdot g = -2x^3 + 2x^2yz + 3xy^3 - (-2x \cdot (x^2+yz)) \nonumber \\
&=& -2x^3 + 2x^2yz + 3xy^3 - (-2x^3-2xyz) = 2x^2yz + 3xy^3 + 2xyz
\end{eqnarray}
Notice that the division $f\stackrel{g}{\textstyle\longrightarrow}r$ cancels the leading term of $f$.
\end{Example}

Similarly, we can also define when a polynomial is divided (reduced) by a set of polynomials.
\begin{Definition}
    The {\bf reduction} of a polynomial $f$, by another polynomial $g$, to
    a reduced polynomial $r$ is denoted:
    \begin{equation*}
        f\stackrel{g}{\textstyle\longrightarrow}_{+}r
    \end{equation*}
    which is the transitive and reflective closure of the relation $f\stackrel{g}{\textstyle\longrightarrow}r$. 
    Reduction is carried out using multivariate, polynomial long division. 
    % The long division is performed according to a term-ordering on polynomials, and the division algorithm 
    %terminates when the leading term of the divisor does not divide any 
    %other term in the dividend.

@@

The above example shows the variety can be infinite, finite (non-empty set) 
or empty. It is interesting to note that since we will be operating over
finite fields $\F_{q}$, and any finite set of points is a variety. Likewise,
any variety over $\F_{q}$ is finite (or empty).
Consider the points $\{(a_1,\dots, a_d): a_1, \dots, a_d \in \F_q\}$
in $\F_q^d$. Any single point is a variety of some polynomial system:
{\it e.g.} $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
varieties are also varieties.

@@

The same ideal may have many different bases.
For instance, it is possible to have different sets of polynomials
$\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
ideal, {\it i.e.} 
$\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
variety depends on the ideal, these sets of polynomials have the same 
solutions.

@@

Ideals and their varieties are a key part of computer-algebra based formal
verification. A given hardware design can be transformed into a set
of polynomials over a field, $F = \{f_1, \ldots, f_s\}$.
This set of polynomials gives the system of equations:
\begin{eqnarray}
f_1 = 0 \nonumber \\
\vdots \nonumber \\
f_s = 0 \nonumber
\end{eqnarray}
Using algebra, it is possible to derive new equations from the original 
system.
The ideal $\langle f_1,\ldots, f_s \rangle$ provides a way of analyzing 
such {\it consequences} of a system of polynomials.

\begin{Example}
Given two equations in $\mathbb{R}[x,y,z]$:
\begin{eqnarray}
x=z+1 \nonumber \\
y=x^2+1 \nonumber 
\end{eqnarray}
we can eliminate $x$ to obtain a new equation:
\begin{equation}
y=(z+1)^2+1=z^2+2z+2 \nonumber 
\end{equation}
Let $f_1, f_2, h \in \mathbb{R}[x,y,z]$ be polynomials based on these 
equations:
\begin{eqnarray}
f_1 &= x-z-1 &= 0 \nonumber \\
f_2 &= y-x^2-1 &= 0 \nonumber \\
h   &= y-z^2-2z-2 &= 0 \nonumber
\end{eqnarray}
If $I$ is the ideal generated by $f_1$ and $f_2$, {\it i.e.} 
$I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
\begin{eqnarray}
g_1 &= x+z+1 \nonumber \\
g_2 &= 1     \nonumber \\
h &= g_1\cdot f_1+g_2\cdot f_2  = y-z^2-2z-2 \nonumber
\end{eqnarray}
where $g_1, g_2 \in \mathbb{R}[x,y,z]$.
Thus, we call $h$ a {\bf member of the ideal} $I$.
\end{Example}

@@

Let $\mathbb{F}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{F}^d$ be a point, and $f \in
\mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ 
if $f(\mathbf{a}) = 0$, {\it i.e.} $\mathbf{a}$ is in the variety of $f$.

@@

As mentioned earlier, different polynomial sets may generate the same 
ideal. Some of these generating sets may be a better representation of 
the ideal, and thus provide more information and insight into the properties 
of the ideal. One such ideal representation is a {\bf \Grobner basis}, which has
a number of important properties that can solve numerous polynomial 
decision questions; the following are used in this dissertation:

\begin{itemize}
\item Check for presence or absence of solutions (varieties)
\item Compute dimension of the varieties
\item Test ideal membership of a polynomial
\item Find projection of varieties by eliminating variables in LEX order
\end{itemize}

@@

\begin{Example}
Consider Example \ref{exp:gbsimple}. Let $f = y^2x - x$ be another
polynomial. Note that $f = yf_1 + f_2$, so $f \in I$. If we divide $f$
by $f_1$ first and then by $f_2$, we will obtain a zero
remainder. However, since the set $\{f_1, f_2\}$ is not a Gr\"{o}bner
basis, we find that the reduction $f
\stackrel{f_2}{\textstyle\longrightarrow} x^2 - x
\stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
{\it i.e.} dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
$I$, $G = \{x^2 - x, yx - y, y^2 - x\}$, dividing $f$ by polynomials
in $G$ in any order will always lead to the zero remainder. Therefore,
one can decide ideal membership unequivocally using the Gr\"{o}bner
basis. 
\end{Example}

@@

\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{$F = \{f_1, \dots, f_s\}$, such that $I=\langle f_1, \dots, f_s\rangle$, and term order $>$}
 \KwOut{$G = \{g_1,\dots ,g_t\}$, a Gr\"{o}bner basis of $I$ }
  $G:= F$\;
  \Repeat{$G = G'$}
  {
  	$G' := G$\;
  	\For{ each pair $\{f_{i}, f_{j}\}, i \neq j$ in $G'$} 
	{
		$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$ \;
		\If{$r \neq 0$}
		{
			$G:= G \cup \{r\}$ \;
		}
	}
   }
\caption {Buchberger's Algorithm}\label{alg:gb}
\end{algorithm}

@@

When computing a \Grobner basis, it's important to note that if $lt(f_i)$ 
and $lt(f_j)$ have no common variables, the $Spoly$ reduction step in 
Buchberger's algorithm will reduce to 0.
\begin{Lemma}
In Buchberger's algorithm, when $gcd(lt(f_i),lt(f_j)) = 0$, the $Spoly$ reduction
$$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$$
will produce $r=0$.
\end{Lemma}

@@

\begin{Definition}\label{def:minigb}
A {\bf minimal Gr\"obner basis} for a polynomial ideal $I$ is a \Grobner basis $G$ for $I$ such that
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$,  $lt(g_{i}) \notin \langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
A {\bf minimal} \Grobner basis is a \Grobner basis such that all polynomials
have a coefficient of $1$ and no leading term of any element in $G$ divides 
another in $G$.
Given a \Grobner basis $G$, a minimal \Grobner basis can be
computed as follows:
\begin{enumerate}
\item Minimize every $g_i \in G$, i.e $g_i=g_i/lc(g_i)$
\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
{\it i.e.} remove every polynomial in $G$ whose leading term is divisible by the leading term 
of some other polynomial in $G$.
\end{enumerate}

@@

\begin{Definition}
Let $H = \{h_1, \ldots, h_t\}$ be a minimal Gr\"obner basis.  Apply
the following reduction process: 
\begin{itemize}
\item $h_1 \stackrel{G_1}{\textstyle\longrightarrow}_+ g_1$, where
  $g_1$ is reduced {\it w.r.t.} $G_1 = \{h_2, \ldots, h_t\}$

\item $h_2 \stackrel{G_2}{\textstyle\longrightarrow}_+ g_2$, where
  $g_2$ is reduced {\it w.r.t.} $G_2 = \{g_1, h_3, \ldots, h_t\}$
\item $h_3 \stackrel{G_3}{\textstyle\longrightarrow}_+ g_3$, where
  $g_3$ is reduced {\it w.r.t.} $G_3 = \{g_1, g_2, h_4, \ldots, h_t\}$

\hspace{0.25in} $\vdots$
\vspace{0.1in}
\item $h_t \stackrel{G_t}{\textstyle\longrightarrow}_+ g_t$, where
  $g_t$ is reduced {\it w.r.t.} $G_t = \{g_1, g_2, g_3, \ldots, g_{t-1}\}$
\end{itemize}
Then $G = \{g_1, \ldots, g_t\}$ is a {\bf reduced Gr\"obner basis.}
\end{Definition}

@@

The high computational complexity of the Gr\"obner basis problem is an important issue
because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
the worst case computation time/space cost of its Gr\"obner basis is doubly exponential while the total degree of 
polynomials in the Gr\"obner basis is bounded by 
$O(2(\frac{1}{2}d^2+d)^{2^{n-1}})$ where $d$ is the degree of the ideal and $n$ is the number of variables
\cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
the polynomial is well restricted rather than arbitrary.
Gao {\it et al.} \cite{gao:qe-gf-gb} proves that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
This provides the possibility to make use of Gr\"obner basis under restricted situations,
which is the theoretic basis of this dissertation.

\Grobner basis computation depends on the $Spoly$ computation, which in turn 
depends on the leading terms of polynomials. Thus, different monomial 
orderings can result in different \Grobner basis computations for the 
same ideal. Computation using a DEGREVLEX ordering tends to be least 
difficult, while lex ordering tends to be computationally complex. However, 
lex ordering used in the computation of \Grobner basis is an {\bf elimination
ordering}; that is, the polynomials contained in the resulting \Grobner basis
have continuously eliminated variables in the ordering. This is the topic of 
elimination theory, which is described in the following sections as well as 
its theoretic basis -- the Nullstellensatz theory.

@@

%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
\begin{Theorem}[Weak Nullstellensatz]
%$\left[\bf{Weak\  Nullstellensatz}\right]$ 
Let $J \subset \overline {\F}[x_1, x_2, \cdots, x_d]$ 
be an ideal satisfying $V(J)=\emptyset$. 
Then $J=\overline {\F}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
\begin{equation}
V(J)=\emptyset\ \iff\ J=\overline {\F}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
\end{equation}
\end{Theorem}

@@

{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
system of multivariate polynomial equations (ideal $J$) has common solutions 
in ${\overline {\F}}^d$. For this purpose, we only need to check if 
the ideal is generated by the unit element, {\it i.e.} $1\in J$. 
This approach can be used to evaluate the feasibility of constraints in 
verification problems.
%%%%%%%%%%%%%%%%%%%%%%%%%strong Nullstellensatz%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
An interesting result is one of {\bf Strong 
Nullstellensatz}.
The strong
Nullstellensatz establishes the correspondence between radical ideals
and varieties.

\begin{Theorem}[Strong Nullstellensatz]
\label{thm:sns}
\cite{gb_book} 
Let $\overline{\F}$ be an algebraically closed field, and let $J$
be an ideal in $\overline{\F}[x_1,\dots, x_d]$. 
Then we have $I(V_{\overline{\F}}(J)) =\sqrt{J}$. 
\end{Theorem}

@@

Here $g_8=Z+A\cdot B$ is the {\bf canonical, word-level polynomial } 
representing the function performed by the multiplier $Z=A\cdot B$.
\end{Example}

\section{Concluding Remarks}
Our approach to word-level abstraction of Galois field arithmetic 
circuits applies concepts of polynomial ideals, varieties, \Grobner basis, 
and abstraction theory to implement verifications on sequential circuits. 
The application of these approaches to sequential circuit verification is described in the following chapters.
